---
title: "Results Section of The Impact of Delivery and Dine-In Services on Yelp Restaurant Ratings"
author: "Simona Hu"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
library(readr)
library(ggplot2)
library(here)
library(stargazer)
```

```{r}
dataset_restaurants <- read_csv(here("data", "restaurants_attributes_extracted.csv"))
```

## Descriptive Statistics

```{r cars}
#Mean and SD
m_stars <- mean(dataset_restaurants$stars, na.rm = TRUE)
sd_stars <- sd(dataset_restaurants$stars, na.rm=TRUE)

m_stars
sd_stars

#Histogram
plot_stars <- ggplot(dataset_restaurants, aes(x = stars)) +
  geom_histogram(binwidth = 0.5, fill = "#A6CEE3", color = "white") +
  labs(
    title = "Distribution of Restaurant Star Ratings",
    x = "Star Rating",
    y = "Number of Restaurants"
  ) +
  theme_minimal()

plot_stars
```

Restaurant star ratings range from 1 to 5, with an average of 3.46 (SD = 0.81), indicating generally positive customer evaluations. Most restaurants are concentrated between ratings of approximately 3 and 4.5 stars, with relatively few low-rated establishments.

Lower price categories dominate the sample, suggesting that the dataset is skewed towards more affordable restaurants. Additionally, both delivery and dine-in scores show sufficient variation, supporting their use as explanatory variables in the regression analysis.

## Effect of delivery options on star ratings

```{r pressure, echo=FALSE}
plot_delivery_stars <- ggplot(dataset_restaurants, aes(x = delivery_score, y = stars)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", color = "black") +
  labs(
    title = "Delivery Score and Star Ratings",
    x = "Delivery Score",
    y = "Star Rating"
  ) +
  scale_y_continuous(limits = c(1, 5)) +
  theme_minimal()

plot_delivery_stars

ggsave(
  filename = here("gen", "output", "plot_delivery_stars.png"),
  plot = plot_delivery_stars,
  width = 7,
  height = 5
)
model_delivery <- lm(stars ~ delivery_score, data = dataset_restaurants)
summary(model_delivery)
```

## Moderating role of price range in the delivery options

```{r pressure, echo=FALSE}

plot_delivery_moderation <- ggplot(
  dataset_restaurants,
  aes(x = delivery_score, y = stars, color = factor(price_range))
) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Moderating Effect of Price Range on Delivery Score and Star Ratings",
    x = "Delivery Score",
    y = "Star Rating",
    color = "Price Range"
  ) +
  scale_y_continuous(limits = c(1, 5)) +
  theme_minimal()

plot_delivery_moderation

ggsave(
  filename = here("gen", "output", "plot_delivery_moderation.png"),
  plot = plot_delivery_moderation,
  width = 7,
  height = 5
)
```

## Effect of dine-in experience on star ratings

```{r pressure, echo=FALSE}
plot_dinein_stars <- ggplot(dataset_restaurants, aes(x = dinein_score, y = stars)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", color = "black") +
  labs(
    title = "Dine-In Score and Star Ratings",
    x = "Dine-In Score",
    y = "Star Rating"
  ) +
  scale_y_continuous(limits = c(1, 5)) +
  theme_minimal()

plot_dinein_stars

ggsave(
  filename = here("gen", "output", "plot_dinein_stars.png"),
  plot = plot_dinein_stars,
  width = 7,
  height = 5
)
model_dinein <- lm(stars ~ dinein_score, data = dataset_restaurants)
summary(model_dinein)
```

## Moderating role of price range in the dine-in options

```{r pressure, echo=FALSE}
model_dinein_mod <- lm(stars ~ dinein_score * price_range, data = dataset_restaurants)
summary(model_dinein_mod)

```

## Summary table of models

```{r pressure, echo=FALSE}
stargazer(
  model_delivery,
  model_moderation,
  model_dinein,
  model_dinein_mod,
  ...
)
```

## 
