---
title: "Merging tip and restaurant"
output: html_document
---

Merging Tip complement counts with cleaned restaurant dataset

**Load required package**

```{r}
library(data.table)
```

Ensure data set_restaurant is treated as a data.table

```{r}
setDT(dataset_restaurants)
setDT(tips) 
```

**Prepare merge key**

Ensuring business_id has the same data type in both data sets

```{r}
dataset_restaurants[, business_id := as.character(business_id)]
tips[, business_id := as.character(business_id)]
```

**Aggregate compliment counts at restaurant level**

sum compliment_count for each business_id (because a restaurant can appear multiple times in the tips data set)

```{r}
tips_agg <- tips[, .(
  tip_compliment_count = sum(compliment_count, na.rm = TRUE)
), by = business_id]
```

**Merge compliment counts into restaurant dataset**

Perform a left join: keep all restaurants from dataset_restaurants and add tip_compliment_count when available

```{r}
restaurants_merged <- merge(
  dataset_restaurants,
  tips_agg,
  by = "business_id",
  all.x = TRUE
)
```

Replace NA values (restuarant without tips) with 0

```{r}
restaurants_merged[is.na(tip_compliment_count), tip_compliment_count := 0]
```

**Validation checks** Check if no restaurants were lost during merge (should be the same)

```{r}
nrow(dataset_restaurants)
nrow(restaurants_merged)
```

Summary statistics of compliment counts

```{r}
summary(restaurants_merged$tip_compliment_count)
```

Count how many restaurants received at least one compliment

```{r}
sum(restaurants_merged$tip_compliment_count > 0)
```
