---
title: "analysis"
output: html_document
---

**Set up chuck**
```{r}

library(ggplot2)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

**Exploring star rating**

```{r}
#Summary statistics
summary(dataset_restaurants$stars)

#Mean and SD
m_stars <- mean(dataset_restaurants$stars, na.rm = TRUE)
sd_stars <- sd(dataset_restaurants$stars, na.rm=TRUE)

m_stars
sd_stars

#Histogram
ggplot(dataset_restaurants, aes(x = stars)) +
  geom_histogram(binwidth = 0.5,
                 fill = "#A6CEE3",
                 color = "white") +
  labs(
    title = "Distribution of Restaurant Star Ratings",
    x = "Star Rating",
    y = "Number of Restaurants"
  ) +
  theme_minimal()
```

The figure shows the distribution of restaurant star ratings. Ratings range from 1 to 5 stars, with most restaurants clustered around ratings between approximately 3 and 4.5 stars. The distribution is slightly left-skewed, indicating that relatively few restaurants receive very low ratings.

The average rating in the data set is 3.46 (SD = 0.81), suggesting that overall customer evaluations are generally positive. Since star rating is the dependent variable in our analysis, understanding its distribution provides important context for interpreting the effect of delivery options, dine-experience and price range in later sections.

**Exploring price range**

```{r}
#Summary statistics 

summary(dataset_restaurants$price_range)

#Histogram

ggplot(dataset_restaurants, aes(x = factor(price_range))) +
  geom_bar(fill = "#A6CEE3") +
  labs(
    title = "Distribution of Restaurant Price Range",
    x = "Price Range",
    y = "Number of Restaurants"
  ) +
  theme_minimal()

```


The figure shows the distribution of restaurants across the four price categories. The majority of restaurants fall within the lower price segments (price levels 1 and 2), with price level 2 being the most common category. In contrast, relatively fewer restaurants are classified as higher-priced establishments (levels 3 and 4), with 4 only representing a very small proportion of the sample. 

This indicates that the data set is skewed toward lower priced restaurants. Since price range serves as a moderating variable in the analysis, the distribution should be considered when interpreting interaction effects in later sections

**Cluster score**
```{r}
#Summary Statistics
summary(dataset_restaurants$delivery_score)
summary(dataset_restaurants$dinein_score)

```

```{r}
#Distribution of the cluster score
ggplot(dataset_restaurants, aes(x = delivery_score)) +
  geom_histogram(binwidth = 1, fill = "#A6CEE3", color = "white") +
  labs(title="Distribution of Delivery Score", x="Delivery Score", y="Number of Restaurants") +
  theme_minimal()

ggplot(dataset_restaurants, aes(x = dinein_score)) +
  geom_histogram(binwidth = 1, fill = "#A6CEE3", color = "white") +
  labs(title="Distribution of Dine-in Score", x="Dine-in Score", y="Number of Restaurants") +
  theme_minimal()

#Scatterplot for Delivery_score
ggplot(dataset_restaurants, aes(x = delivery_score, y = stars)) +
  geom_smooth(method = "lm", color = "black") +
  labs(
    title = "Delivery Score and Star Ratings",
    x = "Delivery Score",
    y = "Star Rating"
  ) +
  scale_y_continuous(limits = c(1,5)) +
  theme_minimal()
```

```{r}
#Delivery score
  model_delivery <- lm(stars ~ delivery_score, data = dataset_restaurants)
#Summary of the delivery score
summary(model_delivery)
```

The delivery and dine-in scores take discrete values, as they are constructed from multiple binary service indicates. The majority of restaurants fall within the lower-to-middle range of both scores, indicating that only a subset of establishments offers many delivery or dine-in related services. 

A simple linear regression analysis shows a statistically significant negative relationship between delivery score and star rating. Restaurants offering more delivery-related services tend to receive slightly lower ratings.

However, the effect size is small. A one-point increase in delivery score is associated with a decrease of only 0.063 stars. Even across the full delivery scale, the maximum predicted difference in rating is approximately 0.25 stars.

Furthermore, delivery score explains less than 1% of the variance in star ratings, indicating that tis pratcial importance is limited.


**Testing Moderation effect with delivery_score**
```{r}
#Moderation effect
model_moderation <- lm(stars ~ delivery_score * price_range, 
                       data = dataset_restaurants)
#Summary of moderation
summary(model_moderation)
```

A regression model including an interaction between delivery score and price range reveals a positive and statistically significant interaction effect (B = 0,168, p < 001). This indicates that price range moderates the relationship between delivery services and star ratings. 

Specifically, the negative association between delivery score and ratings becomes weaker as price range increases and eventually returns positive for higher-priced restaurants. For lower-priced establishments, offering more delivery-related services is associated with slightly lower ratings. In contrast, for higher-priced restaurants, delivery services are associated with higher ratings. 

The inclusion of the interaction improves model fit (R^2 = 0.047), compared to the model without moderation (R^2 = 0.005). This suggests that price range plays a meaningful moderating role in explaining variation in restaurant ratings

```{r}
#Data plot for delivery_score
ggplot(dataset_restaurants,
       aes(x = delivery_score,
           y = stars,
           color = factor(price_range))) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous(limits = c(1,5)) +
  theme_minimal()
```

This figure displays the moderating role of price range on the relationship between delivery score and star ratings. For lower-priced restaurants, an increase in delivery-related services is associated with lower customer ratings, as indicated by the negative slope. As price range increases, this relationship weakens and gradually becomes positive. 

For higher-priced restaurants, delivery services are associated with higher ratings, consistent with the positive interaction effect osberved in the regression analysis. Although the slope of the highest price category appears relatively flat visually, the regression results confirm a statistically significant positive interaction effect, indicating that the association between delivery services and ratings becomes more favorable at higher price levels. 

**Dine_in_score**

```{r}
#summary statistics
summary(dataset_restaurants$dinein_score)

#Dataplot for dine_in_score        
ggplot(dataset_restaurants, aes(x = dinein_score, y = stars)) +
  geom_smooth(method = "lm", color = "black") +
  scale_y_continuous(limits = c(1,5)) +
  labs(
    title = "Dine-in Score and Star Ratings",
    x = "Dine-in Score",
    y = "Star Rating"
  ) +
  theme_minimal()   

```

This figure shows the relationship between dine-in score and star ratings. The fitted regression line indicates a positive association, suggesting that restaurants offering more dine-in related services tend to receive higher customer ratings. The slope appears steeper than that observed for delivery score, indicating that dine-in experience may have a stronger relationship with customer evaluations in this dataset.

```{r}
#Dine_in_score summary
model_dinein <- lm(stars ~ dinein_score,
                   data = dataset_restaurants)
summary(model_dinein)
```
A simple linear regression shows a statistically significant positive relationship between dine-in score and star rating (P <.001). This indicates that restaurants offering  more dine-in related services tend to receive higher customer ratings.

The effect size is meaningful: across the observed range of the dine-in-score, the predicted difference in ratings is approximately 0.58 stars.

The model explains about 2.9% of the variance in star ratings, indicating that dine-in experience is a stronger predictor of ratings than delivery options in this dataset.

**Moderation for dine-in**
```{r}
#Summary for dine_in_moderation
model_dinein_mod <- lm(stars ~ dinein_score * price_range, data = dataset_restaurants)
summary(model_dinein_mod)
```
A regression model including an interaction between dine-in score and price range reveals a statistically significant moderation effect. The interaction indicates that the positive association between dine-in experience and star ratings becomes weaker as price range increases. 

Simple slope calculations suggests that dine-in score is positively related to ratings for lower-priced restaurants. However, the strength of this relationship diminishes for mid-priced establishments and may become slightly negative for the higher price categories.

This pattern suggest that dine-in features are particularly valued in lower-priced restaurants, whereas for higher-priced establishments, additional dine-in services do not necessarily translate into higher customer ratings.

This finding contrasts with the moderation observed for delivery services, suggesting that the strategic importance of dine-in features differs across price segments. 

```{r}
#Data plot for dine_in_score and star ratings by price range
ggplot(dataset_restaurants,
       aes(x = dinein_score, y = stars, color = factor(price_range))) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous(limits = c(1,5)) +
  labs(
    title = "Dine-in Score and Star Ratings by Price Range",
    x = "Dine-in Score",
    y = "Star Rating",
    color = "Price Range"
  ) +
  theme_minimal()
```

This figure illustrates the moderating role of price range on the relationship between dine-in score and star ratings. Across most price levels, dine-in score is positively associated with customer ratings. However, the strength of this relationship diminishes as price range increases. 

For lower-priced restaurants, improvements in dine-in features are associated with relatively larger increases in star ratings. In contrast, for higher priced-establishments, the relationship is weaker and may approach zero.

This pattern is consistent with the negative interaction term observed in the regression analysis. Indicating that the effect of dine_in experience on ratings diminishes at higher price levels 
